Constant-sized correlations are sufficient for robust self-testing of maximally entangled states with unbounded dimension

### 2019-05-05 Draft/draft.tex
Honghao: edits to section 1

At a high level, general $d$-dimensional maximally entangled states are self-tested by
modifying the correlation and enlarging the size of the correlation.
A natural question to ask is whether maximally entangled state with large local dimension
can be self-tested with fixed-sized correlation. 
An equivalent question to ask is whether it is possible to self-test maximally
entangled state of some local dimension more efficiently, with constant correlation size. 
In this report, we give an affirmative answer to this question by proving the following theorem.
\begin{theorem}[Informal]
\label{thm:inf}
	There exists an infinite-sized set $D$ of odd prime numbers such that, for any $d \in D$, 
	the maximally entangled state of local dimension $d-1$ can be self-tested 
	with constant-sized question and answer alphabets.
\end{theorem}

The set $D$ is easily characterizable as it contains all the odd prime numbers with smallest
primitive root $2$, $3$ or $5$. It has been shown that there are infinitely many prime numbers
with smallest primitive root in the set $\{2,3,5\}$ \cite{murty1988}, so the set $D$ has infinitely many elements.
To prove \cref{thm:inf}, we give explicit self-testing proof of 
the maximally entangled state with local dimension $d-1$ where the primitive root of $d$ is $2$,$3$, or $5$, 
by explicitly giving the correlation that achieves self-testing. 
%Then the self-testing property of this correlation is 
%given in the following theorem.
%\begin{theorem}[Informal]
%\label{thm:pr_2}
%	All maximally entangled state with local dimension $d-1$, where $d$ is prime and has
%	primitive root $r \in \{2, 3, 5\}$, can be self-tested by a constant-sized correlation.
%\end{theorem}
Our correlation is denoted by $C(\dr{r})$ for prime $d$.
We use the superscript $(r)$ to denote the primitive root of $d$.
Note that although the size of $C(\dr{r})$ does not depend on $d$, the optimal correlation does.

In order to accomplish our goal, we introduce new techniques for self-testing.
First of all, we use a different variant of the weighted CHSH inequality, \hf{which was first introduced in \cite{acin2012}}, to enforce the eigenvalue of
some unknown operator which is the product of two binary observables used in the weighted CHSH test.
The variant of the CHSH inequality that we use is not used in the self-testing literature before.
Secondly, we give a new way to decompose unitaries of arbitrary order into binary observables which maintains
certain commutation relations. Such decomposition is different from what Slofstra used in his work \cite{slofstra2017}.
\carl{Our work is heavily based on \cite{slofstra2017}, and we need to make that clearer.}
Intuitively, such decomposition can be seen as the inverse of the Jordan's lemma decomposition.
The third contribution is that we prove self-testing without using anti-commutation relations between Pauli operators, 
which is the core idea in \hf{most of} the previous self-testing results.
\carl{Are we sure about that ("all")?}
Instead, we find a new pair of operators that can generate the ring of matrices over complex numbers.

\textbf{Structure of the paper}.
We start with notations and background information in \cref{sec:prelim}.
Since the correlation we designed can win a special linear system game and satisfy
an extended weighted CHSH test, we introduce the linear system game
in \cref{sec:lsg} and the extended weighted CHSH test in \cref{sec:chsh}. 
Our main result is based on the combination of the two tests and presented in \cref{sec:main}.

@@

	For the other relations, we prove $M_{\nr+1}^0M_{r+6} \ket{\psi} \appd{O(r\se)} M_{r+6}M_{\nr+1}^0 \ket{\psi}$ to demonstrate 
	the techniques involved. We denote Bob's projectors by $\{ N_{(\nr+1, r+6)}^{b_1 b_2} \}$ where $b_1 \in \{0, 1, \perp\}$ and
	$b_2 \in \{0,1\}$. From the correlation, we can get that 
	\begin{align*}
		&M_{\nr+1}^0 \ket{\psi} \appd{O(r\se)} (N_{(\nr+1, r+6)}^{00} + N_{(\nr+1, r+6)}^{01}) \ket{\psi} \\
		&M_{r+6} \ket{\psi} \appd{O(r\se)} (\sum_{b_1, b_2} (-1)^{b_2}N_{(\nr+1, r+6)}^{b_1b_2})\ket{\psi}.
	\end{align*}
	Then we can show that
	\begin{align*}
		M_{\nr+1}^0 M_{r+6} \ket{\psi} &\appd{O(r\se)} M_{\nr+1}^0 (\sum_{b_1, b_2} (-1)^{b_2}N_{(\nr+1, r+6)}^{b_1b_2})\ket{\psi}\\
		&\appd{O(r\se)}(\sum_{b_1, b_2} (-1)^{b_2}N_{(\nr+1, r+6)}^{b_1b_2}) (N_{(\nr+1, r+6)}^{00} + N_{(\nr+1, r+6)}^{01})
		\ket{\psi} \\
		&= (N_{(\nr+1, r+6)}^{00} + N_{(\nr+1, r+6)}^{01}) (\sum_{b_1, b_2} (-1)^{b_2}N_{(\nr+1, r+6)}^{b_1b_2})  \ket{\psi} \\
		& \appd{O(r\se)} (N_{(\nr+1, r+6)}^{00} + N_{(\nr+1, r+6)}^{01}) M_{r+6} \ket{\psi} \\
		&= M_{r+6}  (N_{(\nr+1, r+6)}^{00} + N_{(\nr+1, r+6)}^{01}) \ket{\psi} \\
		& \appd{O(r\se)} M_{r+6}M_{\nr+1}^0 \ket{\psi}.
	\end{align*}
\end{proof}
%-----------------------------------------------
\subsection{Robust self-testing}
\label{sec:self-test}
%-----------------------------------------------
With the previous propositions and lemmas, we can prove the self-testing theorem.
\begin{theorem}
	\label{thm:self-test}
	Let $S$ be an $\ep$-approximate strategy of $\fC(d,r)$.
	There exists an isometry   
	$\Phi_A \x \Phi_B = (\Phi_{A,2} \x \Phi_{B,2})(\Phi_{A,1}\x\Phi_{B,2})$ and a state $\ket{junk}$
	such that $\norm{\ket{junk}}^2 \appd{O(dr\se)} 1$ and 
	\begin{align}
		&\Phi_A \x \Phi_B (\ket{\psi}) \appd{O(d^{5/2} r^{d} \ep^{1/8}))} \ket{junk} \x \ket{\EPR{2}}^{\x 2} \x \ket{\EPR{d-1}}\\	
		&\Phi_A \x \Phi_B (O_A\ket{\psi}) \appd{O(d^{5/2} r^{d}\ep^{1/8}))} \ket{junk} \x \ket{\EPR{2}}^{\x 2} \x
		\frac{1}{\sqrt{d-1}}\sum_{j=1}^{d-1} \omega_d^{d-j}\ket{x_{d-j}}_{A'}\ket{x_j}_{B'} \\
		&\Phi_A \x \Phi_B (O_B\ket{\psi}) \appd{O(d^{5/2}r^{d} \ep^{1/8}))} \ket{junk} \x \ket{\EPR{2}}^{\x 2} \x
		\frac{1}{\sqrt{d-1}}\sum_{j=1}^{d-1} \omega_d^{j}\ket{x_{d-j}}_{A'}\ket{x_j}_{B'}\\
		&\Phi_A \x \Phi_B (U_A\ket{\psi}) \appd{O(d^{5/2} r^{d}  \ep^{1/8}))} \ket{junk} \x \ket{\EPR{2}}^{\x 2} \x
		\frac{1}{\sqrt{d-1}}\sum_{j=1}^{d-1} \ket{x_{(d-j)r^{-1}}}_{A'}\ket{x_j}_{B'} \\
		&\Phi_A \x \Phi_B (U_B\ket{\psi}) \appd{O(d^{5/2} r^{d}\ep^{1/8}))} \ket{junk} \x \ket{\EPR{2}}^{\x 2} \x
		\frac{1}{\sqrt{d-1}}\sum_{j=1}^{d-1} \ket{x_{d-j}}_{A'}\ket{x_{j r^{-1}}}_{B'}.
	\end{align}
\end{theorem}
Before proving \cref{thm:self-test}, we state the consequence of it.
It implies that,
for any odd prime number $d$ with primitive root $r$, 
there exists a correlation of size $\Theta(r^2)$ that can robustly self-test the maximally
entangled state of local dimension $4(d-1)$.
Moreover, there are infinitely many prime numbers
whose primitive roots are in the set $\{2,3,5\}$ \cite{murty1988},
Hence, our main result is the following theorem.
\begin{theorem}
	There exists an infinity-sized set $D$ of prime numbers such that 
	for each $d \in D$, there exists
	a constant-sized correlation that can self-test the maximally entangled state
	of local dimension $4(d-1)$.
\end{theorem}

>>>>

A natural question to ask is whether maximally entangled states with large local dimension
can be self-tested with constant-sized correlations. 
In this paper, we give an affirmative answer to this question by proving the following theorem.
\begin{theorem}[Informal]
\label{thm:inf}
	There exists an infinite-sized set $D$ of odd prime numbers such that, for any $d \in D$, 
	the maximally entangled state of local dimension $4d-4$ can be robustly self-tested 
	with constant-sized question and answer alphabets.
\end{theorem}
The formal statement of this theorem is given in \cref{thm:self-test} and \cref{thm:infty}.

The set $D$ is easily characterizable as it contains all the odd prime numbers with smallest
primitive root $2$, $3$ or $5$. It has been shown that  the set $D$ has infinitely many elements \cite{murty1988}.
To prove \cref{thm:inf}, we construct a correlation of size $\Theta(r^2)$ for each $d \in D$ whose primitive
root is $r$. This correlation is denoted by $\fC(d,r)$.
Note that although the size of $\fC(d,r)$ does not depend on $d$, the entries of $\fC(d,r)$ do.
Then we prove the robust self-testing property of $\fC(d,r)$ by showing 
that if some quantum strategy can approximate $\fC(d,r)$, then
the generalized swap isometry \cite{yang2013}
can produce the desired maximally entangled states when applied to the unknown state used in the strategy.
%Then the self-testing property of this correlation is 
%given in the following theorem.
%\begin{theorem}[Informal]
%\label{thm:pr_2}
%	All maximally entangled state with local dimension $d-1$, where $d$ is prime and has
%	primitive root $r \in \{2, 3, 5\}$, can be self-tested by a constant-sized correlation.
%\end{theorem}

In order to accomplish our goal, we introduce new techniques for robust self-testing.
We use a different variant of the weighted CHSH inequality, \hf{which was first introduced in Ref.~\cite{acin2012}
for studying randomness and entanglement and not used in the self-testing literature before}, 
to enforce the eigenvalue of the product of two binary observables used in the weighted CHSH test.
The idea of decompose the unitaries of arbitrary order into binary observables was first introduced by Slofstra \cite{slofstra2017}.
We use a different way to decompose the unitaries, which can be seen as the inverse of Jordan's lemma, and
apply it to self-testing proofs.
Comparing to the previous approaches, which aim to enforcing eigenvalues of individual operators, our approach
gives us more flexibility in designing the correlation and easier generalization of our proof techniques. 
The other contribution is that we prove self-testing without using the anti-commutation relation between Pauli operators,
or the generalized version of it, 
which is the core idea in \hf{most of} the previous self-testing results.
\carl{Are we sure about that ("all")?}
Instead, we find a new relation that should be enforced.

pair of operators that can generate the ring of matrices over complex numbers.
\carl{Our work is heavily based on \cite{slofstra2017}, and we need to make that clearer.}
\hfc{How about now?}

\textbf{Structure of the paper}.
We start with notations and background information in \cref{sec:prelim}.
Since the correlation we designed can win a special linear system game and satisfy
an extended weighted CHSH test, we introduce the linear system game
in \cref{sec:lsg} and the self-testing properties of the extended weighted CHSH inequality in \cref{sec:chsh}. 
Then we give the correlation $\fC(d,r)$ and the ideal strategy to generate it in \cref{sec:full_test}.
In the end, we give the robust self-testing proof in \cref{sec:main}.

@@

	For the other relations, we prove $M_{\nr+1}^0M_{r+6} \ket{\psi} \appd{O(r\se)} M_{r+6}M_{\nr+1}^0 \ket{\psi}$ to demonstrate 
	the techniques involved. We denote Bob's projectors by $\{ N_{(\nr+1, r+6)}^{b_1 b_2} \}$ where $b_1 \in \{0, 1, \perp\}$ and
	$b_2 \in \{0,1\}$. From the correlation, we can get that 
	\begin{align*}
		&M_{\nr+1}^0 \ket{\psi} \appd{O(r\se)} (N_{(\nr+1, r+6)}^{00} + N_{(\nr+1, r+6)}^{01}) \ket{\psi} \\
		&M_{r+6} \ket{\psi} \appd{O(r\se)} (\sum_{b_1, b_2} (-1)^{b_2}N_{(\nr+1, r+6)}^{b_1b_2})\ket{\psi}.
	\end{align*}
	Then we can show that
	\begin{align*}
		M_{\nr+1}^0 M_{r+6} \ket{\psi} &\appd{O(r\se)} M_{\nr+1}^0 (\sum_{b_1, b_2} (-1)^{b_2}N_{(\nr+1, r+6)}^{b_1b_2})\ket{\psi}\\
		&\appd{O(r\se)}(\sum_{b_1, b_2} (-1)^{b_2}N_{(\nr+1, r+6)}^{b_1b_2}) (N_{(\nr+1, r+6)}^{00} + N_{(\nr+1, r+6)}^{01})
		\ket{\psi} \\
		&= (N_{(\nr+1, r+6)}^{00} + N_{(\nr+1, r+6)}^{01}) (\sum_{b_1, b_2} (-1)^{b_2}N_{(\nr+1, r+6)}^{b_1b_2})  \ket{\psi} \\
		& \appd{O(r\se)} (N_{(\nr+1, r+6)}^{00} + N_{(\nr+1, r+6)}^{01}) M_{r+6} \ket{\psi} \\
		&= M_{r+6}  (N_{(\nr+1, r+6)}^{00} + N_{(\nr+1, r+6)}^{01}) \ket{\psi} \\
		& \appd{O(r\se)} M_{r+6}M_{\nr+1}^0 \ket{\psi}.
	\end{align*}
\end{proof}
%-----------------------------------------------
\subsection{Robust self-testing}
\label{sec:self-test}
%-----------------------------------------------
With the previous propositions and lemmas, we can prove the self-testing theorem.
\begin{theorem}
	\label{thm:self-test}
	Let $S$ be an $\ep$-approximate strategy of $\fC(d,r)$.
	There exists an isometry   
	$\Phi_A \x \Phi_B = (\Phi_{A,2} \x \Phi_{B,2})(\Phi_{A,1}\x\Phi_{B,2})$ and a state $\ket{junk}$
	such that $\norm{\ket{junk}}^2 \appd{O(dr\se)} 1$ and 
	\begin{align}
		&\Phi_A \x \Phi_B (\ket{\psi}) \appd{O(d^{5/2} r^{d} \ep^{1/8}))} \ket{junk} \x \ket{\EPR{2}}^{\x 2} \x \ket{\EPR{d-1}}\\	
		&\Phi_A \x \Phi_B (O_A\ket{\psi}) \appd{O(d^{5/2} r^{d}\ep^{1/8}))} \ket{junk} \x \ket{\EPR{2}}^{\x 2} \x
		\frac{1}{\sqrt{d-1}}\sum_{j=1}^{d-1} \omega_d^{d-j}\ket{x_{d-j}}_{A'}\ket{x_j}_{B'} \\
		&\Phi_A \x \Phi_B (O_B\ket{\psi}) \appd{O(d^{5/2}r^{d} \ep^{1/8}))} \ket{junk} \x \ket{\EPR{2}}^{\x 2} \x
		\frac{1}{\sqrt{d-1}}\sum_{j=1}^{d-1} \omega_d^{j}\ket{x_{d-j}}_{A'}\ket{x_j}_{B'}\\
		&\Phi_A \x \Phi_B (U_A\ket{\psi}) \appd{O(d^{5/2} r^{d}  \ep^{1/8}))} \ket{junk} \x \ket{\EPR{2}}^{\x 2} \x
		\frac{1}{\sqrt{d-1}}\sum_{j=1}^{d-1} \ket{x_{(d-j)r^{-1}}}_{A'}\ket{x_j}_{B'} \\
		&\Phi_A \x \Phi_B (U_B\ket{\psi}) \appd{O(d^{5/2} r^{d}\ep^{1/8}))} \ket{junk} \x \ket{\EPR{2}}^{\x 2} \x
		\frac{1}{\sqrt{d-1}}\sum_{j=1}^{d-1} \ket{x_{d-j}}_{A'}\ket{x_{j r^{-1}}}_{B'}.
	\end{align}
\end{theorem}
Before proving \cref{thm:self-test}, we state the consequence of it.
It implies that,
for any odd prime number $d$ with primitive root $r$, 
there exists a correlation of size $\Theta(r^2)$ that can robustly self-test the maximally
entangled state of local dimension $4(d-1)$.
Moreover, there are infinitely many prime numbers
whose primitive roots are in the set $\{2,3,5\}$ \cite{murty1988},
Hence, our main result is the following theorem.
\begin{theorem}
	\label{thm:infty}
	There exists an infinity-sized set $D$ of prime numbers such that 
	for each $d \in D$, there exists
	a constant-sized correlation that can self-test the maximally entangled state
	of local dimension $4(d-1)$.
\end{theorem}
